---
title: "Reproducible Research: Peer Assessment 1"
author: Carlos Toro
output: 
  html_document:
    keep_md: true
---

### Global Options

The Global options are set to show all R code used to make this markdown.

```{r setoptions}
library(knitr)
opts_chunk$set(echo=TRUE)
```

### Loading and preprocessing the data

1. First, the working directory is set, then the data is loaded and assigned to **activityData** variable by using *read.csv()* function.
```{r loaddata}
setwd("/Users/usuario/Desktop/COURSERA/DSC/5. Reproducible Research/Project1")
activityData <- read.csv("activity.csv", colClasses=c("numeric", "Date", "numeric"))
```

2. The variables types are changed to ease the manipulation in R. **steps** and **internal** are converted to R *numeric* type and **date** is converted to R *date* type. A summary of the data is shown below:

```{r summarydata}
summary(activityData)
```

### What is mean total number of steps taken per day?

1. This is the Histogram of the total number of steps taken each day:
```{r histogramSteps}
totalSteps <- aggregate(steps ~ date, activityData, sum)
hist(totalSteps$steps, main="Total number of steps taken per day", xlab = "Total steps", ylab = "Count", col = "blue", breaks = 11)
```

2. The mean and median of the total number of steps taken per day is calculated:
```{r meanAndMedian}
meanSteps <- mean(totalSteps$steps)
meanSteps
medianSteps <- median(totalSteps$steps)
medianSteps
```
  a. The mean of total number of steps taken per day is *`r meanSteps`* steps.
  b. The median of total number of steps taken per day is *`r medianSteps`* steps.

### What is the average daily activity pattern?

1. The time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):
```{r timeseriesPlot}
avgDailysteps <- aggregate(steps ~ interval, activityData, mean)
plot(avgDailysteps, type="l", xlab="Time Intervals (5-minute)", ylab="Average number of steps taken (all Days)", main = "Average steps taken accross all days", col="blue")
```

2. The maximum number of steps is calculates as its contained on the interval on average across all the days in the dataset. 

```{r maxinterval}
maxInterval <- avgDailysteps$interval[which.max(avgDailysteps$steps)]
maxInterval
```
is *`r maxInterval`th* minute.

### Imputing missing values

1. The total number of rows with missing values (NAs) in the dataset 
```{r totalNA}
totalNA <- sum(is.na(activityData))
totalNA
```
is *`r totalNA`*.

2. The mean value across all days is replaced with all the missing values (NAs) in the dataset.
```{r imputing}
activityDataNA <- activityData
activityDataNA$steps[is.na(activityDataNA$steps)] <- avgDailysteps$steps
```

3. The new dataset **activityDataNA** is the same as the original dataset **activityData** with the difference that the missing values of the original were filled in with the mean values across all days.

```{r newdataset}
summary(activityData)
summary(activityDataNA)
```

4. This is the histogram of the total number of steps taken each day with the missing values replaced with the average:
```{r histogramNA}
totalStepsNA <- aggregate(steps ~ date, activityDataNA, sum)
hist(totalStepsNA$steps, main="Total number of steps taken per day", xlab = "Total steps", ylab = "Count", col = "blue", breaks = 11)
```

The mean and median of the total number of steps taken per day with the missing values replaced with the average:
```{r meanAndMedianNA}
meanStepsNA <- mean(totalStepsNA$steps)
meanStepsNA
medianStepsNA <- median(totalStepsNA$steps)
medianStepsNA
```
a. The mean of the total number of steps taken per day is *`r meanStepsNA`* steps.
b. The median of the total number of steps taken per day is *`r medianStepsNA`* steps.

### Are there differences in activity patterns between weekdays and weekends?

1. First we sepaarated the data to see wether its a Weekday or Weekend, as indicated on the date.
```{r dayType}
weekActivity <- activityData
weekActivity$weekdays <- factor(format(weekActivity$date, "%A"))
levels(weekActivity$weekdays) <- list(weekday=c("lunes", "martes", "miércoles", "jueves", "viernes"), weekend = c("sábado", "domingo"))

```

2. The time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days and all the weekend days (y-axis):

```{r weekdaysVsweekends}
avgWeekSteps <- aggregate(steps ~ interval + weekdays, weekActivity, "mean")

library(lattice)
xyplot(steps ~ interval | weekdays, avgWeekSteps, layout=c(1, 2), type="l", xlab="Interval (5-minute)", ylab="Average number of steps across all days")

```

